### 2020-04-23 Case Studies/02_race-times.rmd
cherry blossom analysis
cherry blossom analysis

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

@@

file2001 <- read_results("men", 2001)
age2001 <- as.numeric(extractVariables(file2001)[, "ag"])

@@

file <- read_results("men", 2001)
data <- extractVariables(file)
age <- as.numeric(data[, "ag"])

@@

```{r}
file2002 <- read_results("men", 2012)

@@

file2006 <- read_results("men", 2006)
parsed2006 <- extractVariables(file2006)
time2006 <- parsed2006[, "net"]

@@

@@

```{r}
mR.models <- cbMenSub %>% 
   group_by(year) %>%
   group_map(~ loess(runTime ~ age, .x))

mR.pred <- lapply(mR.models, 
                  function(m, year, age) cbind(rep(year, length(age)), age = age, predict(m, data.frame(age = age))),
                  year = 1999:2012,
                  age = age20to80)

yoy_changes <- matrix(unlist(mR.pred), ncol = 3)
yoy_changes <- as.data.frame(yoy_changes)
colnames(yoy_changes) <- c("year", "pred", "age")
yoy_changes$year <- as.numeric(yoy_changes$year)

ggplot(yoy_changes[year == 1999], aes(age, pred)) +
   geom_point()
```

>>>>

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 4.5)

@@

file2001 <- load_results("men", 2001)
age2001 <- as.numeric(extractVariables(file2001)[, "ag"])

@@

file <- load_results("men", 2001)
data <- extractVariables(file)
age <- as.numeric(data[, "ag"])

@@

```{r}
file2002 <- load_results("men", 2012)

@@

file2006 <- load_results("men", 2006)
parsed2006 <- extractVariables(file2006)
time2006 <- parsed2006[, "net"]

@@

summary(mR.lo.pr99)

@@

years <- 1999:2012
results <- list(length(years))

y <- 1999

for( i in 1:length(years) )
{
   y = years[i]
   data <- cbMenSub[ which(cbMenSub$year == y), ]

   model <- loess(runTime ~ age, data)

   pred <- predict(model, newdata = age20to80)

   results[[i]] <- data.table(Year = rep(y, length(pred)), Age = age20to80, Pred = pred)
}

race_years_data <- do.call("rbind", results)

ggplot(race_years_data, aes(Age, Pred, group = Year)) +
   geom_line(aes(col = Year)) +
   labs(title = "Predicted Performance by Year", ylab = "Predicted Run Time", xlab = "Runner Age")

### Constructing a Record for an Individual Runner across Years

```{r}
trimBlanks <- function(charVector) {
   nameClean <- gsub("^[[:blank:]]+", "", charVector)
   nameClean <- gsub("[[:blank:]]+$", "", nameClean)
   nameClean <- gsub("[[:blank:]]+", " ", nameClean)
}

nameClean <- trimBlanks(cbMenSub$name)

length(nameClean)

length(unique(nameClean))

table(table(nameClean))

```{r}
head(sort(table(nameClean), decreasing = T), 1)

mSmith <- cbMenSub[nameClean == "Michael Smith", ]

head(unique(mSmith$home))

nameClean <- tolower(nameClean)

head(sort(table(nameClean), decreasing = T), 1)

nameClean <- gsub("[,.]", "", nameClean)
```

```{r}
tabNameYr <- table(cbMenSub$year, nameClean)

max(tabNameYr)

class(tabNameYr)

mode(tabNameYr)

names(attributes(tabNameYr))

dim(tabNameYr)

head(colnames(tabNameYr), 3)

which(tabNameYr == max(tabNameYr), arr.ind = T)

indMax <- which(tabNameYr == max(tabNameYr), arr.ind = T)
colnames(tabNameYr)[indMax[2]]

cbMenSub$nameClean <- nameClean
```

```{r}
cbMenSub$yob <- cbMenSub$year - cbMenSub$age

homeClean <- trimBlanks(cbMenSub$home)
homeClean <- tolower(homeClean)

cbMenSub$homeClean <- homeClean
```

```{r}
vars <- c("year", "homeClean", "nameClean", "yob", "runTime")
mb <- which(nameClean == "michael brown")
birthOrder <- order(cbMenSub$yob[mb])
cbMenSub[mb[birthOrder], vars]
```

```{r}
cbMenSub$ID = paste(nameClean, cbMenSub$yob, sep = "_")

races <- tapply(cbMenSub$year, cbMenSub$ID, length)

races8 <- names(races)[which(races >= 8)]

men8 <- cbMenSub[ cbMenSub$ID %in% races8, ]

orderByRunner <- order(men8$ID, men8$year)

men8 <- men8[orderByRunner, ]
```

```{r}
men8L <- split(men8, men8$ID)
names(men8L)
```

```{r}
length(unique(men8$ID))
length(men8L)
```

```{r}
gapTime <- tapply(men8$runTime, men8$ID,
                  function(t) any(abs(diff(t)) > 20))

gapTime <- sapply(men8L, function(df)
   any(abs(diff(df$runTime)) > 20))

sum(gapTime)
```

```{r}
lapply(men8L[ gapTime ][1:2], function(df) df[, vars])
```

```{r}
homeLen <- nchar(cbMenSub$homeClean)

cbMenSub$state <- substr(cbMenSub$homeClean,
                         start = homeLen - 1, stop = homeLen)

cbMenSub$state[cbMenSub$year == 2006] = NA

cbMenSub$ID = paste(cbMenSub$nameClean, cbMenSub$yob,
                    cbMenSub$state, sep = "_")
```

```{r}
numRaces <- tapply(cbMenSub$year, cbMenSub$ID, length)
races8 <- names(numRaces)[which(numRaces >= 8)]
men8 <- cbMenSub[ cbMenSub$ID %in% races8, ]

orderByRunner <- order(men8$ID, men8$year)
men8 <- men8[orderByRunner, ]

men8L <- split(men8, men8$ID)
names(men8L) <- races8

length(men8L)
```

### Modeling the Change in Running Time for Individuals
